Mapping mutational effects along the evolutionary landscape of HIV envelope

### 2018-02-26 paper/manuscript.tex
small tweaks to wording on clusters of shifted sites

Indeed, many of the shifted sites cluster within regions of Env that are highly conformationally dynamic.
\FIGSUPP[shiftfeatures]{dynamic_clusters} shows the structural context of these clusters in finer detail.
One cluster is at the trimer apex where two of Env's variable loops pack against one another and against an adjacent protomer.
Upon CD4 binding, these loops become highly disordered~\citep{guttman2014cd4,ozorowski2017open}.
Mutations at two of the shifted sites in this cluster (165 and 307) have been shown to cause Env to assume aberrant conformations, suggesting that these sites can strongly modulate Env's dynamics~\citep{lee2017broadly}.
The other cluster of shifted sites is near a network of hydrophobic amino acids that has been proposed to help transmit the large-scale conformational change that takes place upon CD4 binding~\citep{ozorowski2017open}.
One of the shifted sites (site 69) overlaps with this network, and mutations at another (site 64) have been shown to strongly modulate the relative stability of the open and closed conformations~\citep{de2015immunogenicity}.
In total, these two clusters consist of nearly half of the shifted sites (13 out of 30).
One hypothesis why so many shifted sites cluster in these regions is that their dynamic nature allows long-range epistatic interactions to be readily propagated between substituted sites and distant shifted sites.
Due to the complexity of Env's dynamics, it is difficult to discern exactly how these interactions might occur.
But there is certainly a trend for sites that are conformationally dynamic to also be sites that show shifts in their amino-acid preferences during evolution.

>>>>

Indeed, many of the shifted sites cluster within regions of Env that are highly conformationally dynamic.
\FIGSUPP[shiftfeatures]{dynamic_clusters} shows the structural context of these clusters in finer detail.
One cluster is at the trimer apex where two of Env's variable loops pack against one another and against an adjacent protomer.
These interactions are likely involved in regulating the transition between conformational states, and upon CD4 binding, these loops become highly disordered~\citep{guttman2014cd4,ozorowski2017open}.
Mutations at two of the shifted sites in this cluster (165 and 307) have been shown to cause Env to assume aberrant conformations, suggesting that these sites can strongly modulate Env's dynamics~\citep{lee2017broadly}.
Strikingly, this cluster of shifted sites may may reflect previously observed differences in the conformational dynamics of this regions between these two Envs; the V2 region of BF520 SOSIP trimer is more accessible to deuterium exchange than the BG505 SOSIP trimer~\citep{verkerke2016epitope}.
The other cluster of shifted sites is near a network of hydrophobic amino acids that has been proposed to help transmit the large-scale conformational change that takes place upon CD4 binding~\citep{ozorowski2017open}.
One of the shifted sites (site 69) overlaps with this network, and mutations at another (site 64) have been shown to strongly modulate the relative stability of the open and closed conformations~\citep{de2015immunogenicity}.
In total, these two clusters consist of nearly half of the shifted sites (13 out of 30).
One hypothesis why so many shifted sites cluster in these regions is that their dynamic nature allows long-range epistatic interactions to be readily propagated between substituted sites and distant shifted sites.
It is difficult to discern exactly how these interactions might occur, but there is certainly a trend for sites that are conformationally dynamic to also be sites that show shifts in their amino-acid preferences during evolution.
